---
title: "2018 Exploring"
output: html_notebook
---

## Loading Data
```{r}
library(tidyverse)

teams = read_csv('data/2018/WDataFiles/WTeams.csv')

games = rbind(
  read_csv('data/2018/WDataFiles/WRegularSeasonCompactResults.csv') %>% mutate(Season_Type = 'Regular'),
  read_csv('data/2018/WDataFiles/WNCAATourneyCompactResults.csv') %>% mutate(Season_Type = 'Tournament')
) %>% mutate(game_id = 1:n())  

# Function to flip locations for the loosing team:

flip_location = Vectorize(
  function(loc) {
    new_loc = switch(loc,
                        H = 'A',
                        A = 'H',
                        N = 'N',
                        'Error!')
    return(new_loc)
  }
)

annotate_location = Vectorize(
  function(loc) {
    new_loc = switch(loc,
                        H = 'Home',
                        A = 'Away',
                        N = 'Neutral',
                        'Error!')
    return(new_loc)
  }
)

performances = rbind(
  # Winner records
  games %>% select(TeamID=WTeamID,
                   Score=WScore,
                   Opponent_Score=LScore,
                   Season,
                   DayNum,
                   Location = WLoc,
                   NumOT, 
                   Season_Type, 
                   game_id) %>% 
    mutate(Win=TRUE),
  # Looser records
  games %>% select(TeamID=LTeamID,
                   Score=LScore,
                   Opponent_Score=WScore,
                   Season,
                   DayNum,
                   Location = WLoc,
                   NumOT, 
                   Season_Type, 
                   game_id) %>% 
    mutate(Win=FALSE,
           Location = flip_location(Location))
) %>% mutate(Location = annotate_location(Location))


```


## Homefield advantage

Let's see how a home game advatage might affect the scores:

Since our `performace` dataset is symetrical (2 records per game), lets just look at the winners.

First we look at the difference in scores for winners by location

```{r}
score_dist_plot = performances %>% 
                    filter(Win) %>% 
                    mutate(score_diff = Score - Opponent_Score) %>% 
                    ggplot(aes(x = score_diff, fill = Location)) 

score_dist_plot + geom_density(alpha = 0.5)

score_dist_plot + geom_density(alpha = 0.5) + facet_wrap(~Location, nrow = 3)

```

We see that blowouts are more likely for home games, whereas tight margins of victory seem to be more common in away games. 

```{r}
games %>% 
  filter(WLoc %in% c('H','A')) %>% # Only get home and away games
  pull(WLoc) %>% table() %>% prop.table()
```
supprisingly, 62% of home games are won.

Although most NCAA games are not played at home or away, we will have to account for this signal in order to get a better sense for how teams might play on neutral teritory. 

## Dynasties (long term success)

```{r}
performances %>% 
  filter(Season_Type=='Regular') %>%
  group_by(TeamID, Season) %>% 
  summarise(Scored=mean(Score),
            Allowed=mean(Opponent_Score)) %>% 
  ungroup() %>% 
  group_by(TeamID) %>% 
  nest() %>%
  sample_n(6) %>% 
  unnest() %>%
  left_join(teams) %>%
  ggplot(aes(x=Season,group=TeamName))+
  geom_line(aes(y=Scored,col="avg Scored Points"))+
  geom_line(aes(y=Allowed,col="avg Allowed Points"))+
  facet_wrap(~TeamName,ncol=2)+
  labs(title="Example Plot using dplyr and ggplot",y="Points",x="Season")+
  theme_light()
```

